---
title: "Ceyx Mapping"
format:
  html:
    code-fold: show
    code-tools: true
toc: true
toc-title: Document Contents
number-sections: true
embed-resources: true
---

## read in data and load packages
```{r}
#| output: false
library(vcfR) #v1.14.0
library(ggplot2) #v3.4.4
library(maps) #v3.4.1.1
library(sf) #v1.0-14
library(RColorBrewer) #v1.1-3
#read vcf
vcfR<-read.vcfR("~/Desktop/nmel.ceyx.rad/filtered.snps.vcf.gz")
#read in sampling details
sample.info<-read.csv("~/Desktop/nmel.ceyx.rad/ceyx.full.sampling.csv")
```

## prepare datframes for plotting
```{r}
#prune sampling so it matches the files retained in the vcf
samps<-sample.info[sample.info$ID %in% colnames(vcfR@gt)[-1],]
#round lat and longs to group sampling points within .1 degrees
samps$lat<-round(samps$lat,1)
samps$long<-round(samps$long,1)
#manually combine nearby sampling sites
samps[69,8]<- -10.6
samps[63:65,8]<- -9.5
samps[63:65,9]<- 160
samps[55,8]<- -7.1
samps[55,9]<- 157
samps[1,8]<- -6.9
samps[13,9]<- 145
samps[16,8]<- -6.9
samps[14,8]<- -4.5
samps[14,9]<- 145
samps[107,8]<- -2.1
samps[107,9]<- 147
samps[33:35,8]<- -5.6
samps[33:35,9]<- 150.6
samps[41,8]<- -6.2
samps[41,9]<- 155.3

#split df by species
spec.dfs<-split(samps, samps$species)

#init sampling.df which will be a df of samples grouped by unique lat/long
sampling.df<-data.frame(NULL)
for (i in names(spec.dfs)){
  samps<-spec.dfs[[i]] %>% dplyr::group_by(lat, long) %>% dplyr::summarize(count=dplyr::n())
  df<-cbind(rep(i, times=nrow(samps)), samps)
  sampling.df<-as.data.frame(rbind(sampling.df, df))
}

#fix colnames
colnames(sampling.df)<-c("species","lat","long","count")
#reorder factors
sampling.df$species <- factor(sampling.df$species, levels = c("margarethae","dispar","gentianus","malaitae","meeki","nigromaxilla","collectoris","sacerdotis","solitarius","mulcatus"))
```

## plot just margarethae for the inset
```{r}
#make map#
pac<-map_data("world")

# plot just marg
ggplot()+
  geom_polygon(data = pac, aes(x=long, y = lat, group = group), fill="lightgrey", col="black", cex=.1)+
  coord_sf(xlim = c(115, 130), ylim = c(4, 20)) + 
  geom_point(data = sampling.df[sampling.df$species == "margarethae",], aes(x = long, y = lat, fill=species, size=count), pch=21, color="black", alpha =.8, show.legend=TRUE) +
  theme_classic()+
  scale_fill_manual(values=c("white"))+
  scale_size_continuous(range = c(3,6))+
  guides(fill = guide_legend(override.aes = list(size = 4), order=1, label.theme = element_text(face = "italic")),
         linewidth = guide_legend(order = 2))+
  theme(legend.position = NULL)
ggsave("~/Desktop/nmel.ceyx.rad/margarethae.map.pdf",width = 4, height = 4.5, units="in")
```

## plot all of the nmel species together
We will use Paul Tol's [muted qualitative color scheme](https://personal.sron.nl/~pault/data/colourschemes.pdf), which is a color-blind friendly diverging palette, plus the addition of white, for the outgroup margarethae.
```{r}
# plot all samples
ggplot()+
  geom_polygon(data = pac, aes(x=long, y = lat, group = group), fill="lightgrey", col="black", cex=.1)+
  coord_sf(xlim = c(143, 162), ylim = c(-12, -2)) + 
  geom_point(data = sampling.df, aes(x = long, y = lat, fill=species, size=count), pch=21, color="black", alpha =.8, show.legend=TRUE) +
  theme_classic()+
  scale_fill_manual(values=c("#FFFFFF","#332288","#88CCEE","#44AA99","#117733","#999933","#DDCC77","#CC6677","#882255","#AA4499"))+
  scale_size_continuous(range = c(3,6))+
  guides(fill = guide_legend(override.aes = list(size = 6), order=1, label.theme = element_text(face = "italic")),
         size = guide_legend(order = 2))+
  theme(legend.position = NULL)
ggsave("~/Desktop/nmel.ceyx.rad/sampling.map.pdf",width = 9, height = 4.5, units="in")
```

## plot a blown up inset showing the sampling within the New Georgia group
```{r}
ggplot()+
  geom_polygon(data = pac, aes(x=long, y = lat, group = group), fill="lightgrey", col="black", cex=.1)+
  coord_sf(xlim = c(156.4, 158.3), ylim = c(-7.5, -9)) + 
  geom_point(data = sampling.df, aes(x = long, y = lat, fill=species, size=count), pch=21, color="black", alpha =.8, show.legend=TRUE) +
  theme_classic()+
  scale_fill_manual(values=c("#FFFFFF","#332288","#88CCEE","#44AA99","#117733","#999933","#DDCC77","#CC6677","#882255","#AA4499"))+
  scale_size_continuous(range = c(3,6))+
  guides(fill = guide_legend(override.aes = list(size = 6), order=1, label.theme = element_text(face = "italic")),
         size = guide_legend(order = 2))+
  theme(legend.position = NULL)
ggsave("~/Desktop/nmel.ceyx.rad/NG.sampling.pdf",width = 4, height = 4, units="in")
```

## plot all samples without the NG group dots included
```{r}
# plot all samples without the NG group dots included
ggplot()+
  geom_polygon(data = pac, aes(x=long, y = lat, group = group), fill="lightgrey", col="black", cex=.1)+
  coord_sf(xlim = c(143, 162), ylim = c(-12, -2)) + 
  geom_point(data = sampling.df[sampling.df$species != "collectoris",], aes(x = long, y = lat, fill=species, size=count), pch=21, color="black", alpha =.8, show.legend=TRUE) +
  theme_classic()+
  scale_fill_manual(values=c("#FFFFFF","#332288","#88CCEE","#44AA99","#117733","#999933","#CC6677","#882255","#AA4499"))+
  scale_size_continuous(range = c(3,6))+
  guides(fill = guide_legend(override.aes = list(size = 6), order=1, label.theme = element_text(face = "italic")),
         size = guide_legend(order = 2))+
  theme(legend.position = NULL)
ggsave("~/Desktop/nmel.ceyx.rad/sampling.map.noNG.pdf",width = 9, height = 4.5, units="in")
```

